Modeling Second-Language Learning from a Psychological Perspective

### 2018-03-29 doc/naaclhlt2018-latex/paper.tex
never ends

Despite these promising opportunities, it remains unclear how much of current psychological
theory might be important for tasks such as the Duolingo SLAM challenge.
In the field of education data mining, researchers trying to build
predictive models of student learning have typically relied on traditional, and interpretable, models and approaches
that are rooted in cognitive science~\cite[e.g.,][]{Atkinson:1972rm,Atkinson:1972rz,Corbett1995bkt,Pavlik:2008rm}.
However, a recent paper found that state-of-the-art results could be achieved using deep neural networks with
little or no cognitive theory built in~\cite[so called ``deep knowledge tracing",][]{Piech:2015aa}.
Khajah, Lindsey, \& Mozer (2016) compared deep knowledge tracing (DKT) to more
standard ``Bayesian knowledge tracing" (BKT) models and showed
that it was possible to equate the performance of the the BKT model by additional 
features and parameters that represent core aspects of the psychology of learning
and memory such as forgetting and individual abilities~\cite{Khajah2016hdnt}. An ongoing debate remains in this community 
whether using flexible models with lots of data can improve over more heavily structured, theory-based models \cite{Tang:2016:DNN, xiong2016going, zhang2017incorporating}.

@@

To tune model hyper-parameters and evaluate the usefulness of features, we first
trained the models on the {\tt train} data set and evaluated them on the {\tt dev}
data set. Details of the training sets and the actual files are provided on the
Harvard Dataverse\footnote{https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/8SWHNO}.
Once the model structure was finalized, we trained on the combined
{\tt train} and {\tt dev} data and produced predictions for the {\tt test} data. The
LightGBM hyperparameters used for each model are listed in Table~\ref{lightgbm-params}.

@@

Though not plotted above, we also ran a model lesioning exercise-level features
including client, session type, format, and exercise duration. This model
achieved an AUROC of $.787$, far lower than any other lesion. This points to the
importance of how memory is assess on performance, reflecting insights
from psychology such as the difference between recall and recognition memory
\cite{yonelinas2002nature}.

>>>>

Despite these promising opportunities, it remains unclear how much of current psychological
theory might be important for tasks such as the Duolingo SLAM challenge.
In the field of education data mining, researchers trying to build
predictive models of student learning have typically relied on traditional, and interpretable, models and approaches
that are rooted in cognitive science~\cite[e.g.,][]{Atkinson:1972rm,Atkinson:1972rz,Corbett1995bkt,Pavlik:2008rm}.
However, a recent paper found that state-of-the-art results could be achieved using deep neural networks with
little or no cognitive theory built in~\cite[so called ``deep knowledge tracing",][]{Piech:2015aa}.
Khajah, Lindsey, \& Mozer (2016) compared deep knowledge tracing (DKT) to more
standard ``Bayesian knowledge tracing" (BKT) models and showed
that it was possible to equate the performance of the BKT model by additional 
features and parameters that represent core aspects of the psychology of learning
and memory such as forgetting and individual abilities~\cite{Khajah2016hdnt}. An ongoing debate remains in this community 
whether using flexible models with lots of data can improve over more heavily structured, theory-based models \cite{Tang:2016:DNN, xiong2016going, zhang2017incorporating}.

@@

To tune model hyper-parameters and evaluate the usefulness of features, we first
trained the models on the {\tt train} data set and evaluated them on the {\tt dev}
data set. Details of the datasets and the actual files are provided on the
Harvard Dataverse\footnote{https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/8SWHNO}.
Once the model structure was finalized, we trained on the combined
{\tt train} and {\tt dev} data and produced predictions for the {\tt test} data. The
LightGBM hyperparameters used for each model are listed in Table~\ref{lightgbm-params}.

@@

Though not plotted above, we also ran a model lesioning exercise-level features
including client, session type, format, and exercise duration. This model
achieved an AUROC of $.787$, far lower than any other lesion. This points to the
fact that \emph{how} memory is assessed often includes observed performance 
\cite[e.g., the large literature in psychology on the difference between recall and recognition memory,][]{yonelinas2002nature}.
